Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-02 paper/broom.tex
elabMeth added.
Edits.

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
\usepackage[breaklinks]{hyperref}

@@

\newcommand{\csolve}{{\sc Csolve}\xspace}

@@

@@

\newcommand{\lstml}{\lstset{ %
language=[Objective]Caml, % choose the language of the code
basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={relation, not, : , /\\}
}}
\lstnewenvironment{codeml}
    { % \centering
			\lstml
      \lstset{mathescape=true}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

>>>>

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
\usepackage[breaklinks]{hyperref}
\usepackage[multiple]{footmisc}

@@

\newcommand{\csolve}{{\sc CSolve}\xspace}

@@

\newcommand{\elabMeth}{{\sf elabMeth}}
\newcommand{\frv}{{\sf frv}}
\newcommand{\solve}{{\sf solve}}
\newcommand{\elabClass}{{\sf elabClass}}
\newcommand{\elabCons}{{\sf elabCons}}
\newcommand{\elabClassTable}{{\sf elabClassTable}}

@@

\newcommand{\lstml}{\lstset{ %
language=[Objective]Caml, % choose the language of the code
basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={foreach, do, done, not, : , /\\}
}}
\lstnewenvironment{codeml}
    { % \centering
			\lstml
      \lstset{mathescape=true}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

### 2015-11-02 paper/type-inference.tex
Edits.

We formalize region type inference as an elaboration function from
programs in $\absof{\FB}$ (i.e., region-erased \FB) to programs in
\FB.  The elaboration proceeds in three steps. In the first step we
make use of the observation that region types are \emph{refinements}
of FGJ types with region annotations and, possibly, constraints over
such region annotations, and generate region type templates for
expressions, methods and classes from their FGJ types. The templates
contain region variables ($\rho$) to denote unknown region
annotations, and predicate variables ($\varphi$) to denote unknown
constraints over such region annotations. We also introduce region
variables to denote unknown region arguments in \C{new} expressions,
method calls and function applications. Next, we make use of the
static semantics in Fig.~\ref{fig:fb-staticsem} to build a system of
constraints over region and predicate variables that capture
well-formedness requirements and subtyping relationships between type
templates that must hold for the region elaboration to be valid.
Finally, we solve the constraints by making use of our fixpoint
constraint solving algorithm \csolvestar, which reduces the
constraint solving problem to an abduction problem in a Herbrand
constraint system, and then relies on \csolve, our abduction
solver for that domain. 
% Due to the presence of region-polymorphic
% recursion in \FB, constraints generated by the algorithm can be
% circular. More precisely, constraints generated can assume the form
% $\varphi \Leftrightarrow \phi \wedge F(\varphi)$, where $F$ is a
% non-idempotent substitution function for region variables in
% $\varphi$. In the constraint solving phase, the algorithm then relies
% on a fixpoint constraint solving algorithm called \csolvestar to solve
% the constraints and determine assignments for unknown region and
% predicate variables.

\subsection{Region Type Templates}

@@

The type template for a recursively defined class is computed in two
steps. First, all recursive occurances of the class among the types of
its fields are ignored and the class is templatized as if it is a
non-recursive class. Next, all the recursive occurances are
templatized with respect to the class template computed in the first
step, such that their region annotations are same as the region
parameters of the class. For example, consider a generic
$\C{ListNode}\inang{\C{a}}$ class containing two fields: $\C{data}$ of
type $\C{a}$ and $\C{next}$ of type $\C{ListNode}\inang{\C{a}}$.  where
$\C{a}$ is the type of the data stored in the linked list node.  To
templatize the $\C{ListNode}\inang{a}$ class, we first ignore its
recursive occurance in the type of $\C{next}$ field, and templatize
the type $\C{a}$ of $\C{data}$ field as $\C{a}@\rho_0$. Based on this
type template of $\C{data}$ field, we compute the class's template
as following:
\begin{center}
$\C{ListNode}\inang{\C{a}\extends\ObjZ}
\inang{\rhoalloc_0,\rho_0\,|\,\varphi_0} \extends \ObjZ \{...\}$
\end{center}
Next, we trivially templatize the type of $\C{next}$ field in the body
of the class as $\C{ListNode}\inang{\C{a}}\inang{\rhoalloc_0,\rho_0}$.
The resulting class represents a linked list with spine in the region
$\rhoalloc_0$ and data objects in the region $\rho_0$.

% Observe that our strategy of trivially templatizing the recursive
% occurances of a class type in its definition ensures that no new
% region variables are introduced.

@@

\subsection{Constraint Generation}

Next, we describe our constraint generation algorithm that analyzes a
templatized class definition and generates constraints over region and
predicate variables occuring in the definition, such that the
definition is well-formed if and only if there exists a satisfying
assignment to region and predicate variables.

% Recall that the overarching purpose of region type inference is to
% elaborate programs in $\absof{\FB}$ to region-annotated programs in
% $\FB$. A templatized class definition is only a partially elaborated 
At the heart of our constraint generation algorithm is a function
$\elabExpr$ that elaborates expressions in $\absof{\FB}$ to
region-annotated expressions in $\FB$. Fig.~\ref{fig:fb-elabexpr}
shows the definition of $\elabExpr$ for a subset of expressions in
$\absof{\FB}$, whose corresponding $\FB$ expressions have been
ascribed static semantics in Fig.~\ref{fig:fb-staticsem}. The function
traverses expressions in a syntax-directed manner of a type checker,
introducing fresh region type templates for unknown region types,
while generating constraints that capture the relationship between the
types of various subexpressions and their well-formedness. $\elabExpr$
is defined under the same context as the expression typing judgment in
Fig.~\ref{fig:fb-staticsem} with symbols $\A$,$\ralloc$, and $\env$
retaining their meaning. Functions $\typeOk$ and $\subtypeOk$
(definitions not shown) referred by $\elabExpr$ generate constraints
making use of the syntax-directed well-formedness and subtype
judgments from Fig.~\ref{fig:fb-staticsem}.

The $\elabMeth$ function shown in Fig.~\ref{fig:fb-elabmeth} lifts the
definition of elaboration from expressions to methods. We let methods
to be defined recursively, but for the sake of presentation, we assume
that there is no mutual recursion. This means that there exists a
total order of all the classes in a program, and a total order of all
the method definitions within each class, where definitions occuring
prior in the order do not refer to those that occur later. We assume
that classes and methods are elaborated in that order. Consequently,

\subsection{Constraint Solving}

>>>>

We formalize region type inference as an elaboration function from
programs in $\absof{\FB}$ (i.e., region-erased \FB) to programs in
\FB.  The elaboration proceeds in four steps. In the first step we
make use of the observation that region types are \emph{refinements}
of FGJ types with region annotations and constraints over such region
annotations, and compute polymorphic region type templates for methods
and classes from their FGJ types. The templates contain region
variables ($\rho$) to denote unknown region annotations, and predicate
variables ($\varphi$) to denote unknown constraints over such region
annotations. Free region variables are generalized in types (hence,
polymorphic). Second, we make use of the computed region type
templates to elaborate expressions by introducing region variables to
denote unknown region arguments in \C{new} expressions, method calls
and function applications. While elaborating expressions, we also
build a system of constraints that capture well-formedness
requirements and subtyping relationships between type templates that
must hold (as per the static semantics in Fig.~\ref{fb-staticsem}) for
the elaboration to be valid. Third, we lift expression elaboration and
constraint generation to methods and classes. Finally, we solve the
constraints by making use of our fixpoint constraint solving algorithm
\csolvestar, which reduces the constraint solving problem to an
abduction problem in a Herbrand constraint system, and then relies on
\csolve, our abduction solver for that domain. 
% Due to the presence of region-polymorphic
% recursion in \FB, constraints generated by the algorithm can be
% circular. More precisely, constraints generated can assume the form
% $\varphi \Leftrightarrow \phi \wedge F(\varphi)$, where $F$ is a
% non-idempotent substitution function for region variables in
% $\varphi$. In the constraint solving phase, the algorithm then relies
% on a fixpoint constraint solving algorithm called \csolvestar to solve
% the constraints and determine assignments for unknown region and
% predicate variables.

\subsection{Region Type Templates}
\label{sec:fb-templatization}

@@

The type template for a recursively defined class is computed in two
steps. First, all recursive occurances of the class among the types of
its fields are ignored and the class is templatized as if it is a
non-recursive class. Next, all the recursive occurances are
templatized with respect to the class template computed in the first
step, such that their region annotations are exactly same as the
region parameters of the class. For example, consider a generic
$\C{ListNode}\inang{\C{a}}$ class containing two fields: $\C{data}$ of
type $\C{a}$ and $\C{next}$ of type $\C{ListNode}\inang{\C{a}}$.
where $\C{a}$ is the type of the data stored in the linked list node.
To templatize the $\C{ListNode}\inang{a}$ class, we first ignore its
recursive occurance in the type of $\C{next}$ field, and templatize
the type $\C{a}$ of $\C{data}$ field as $\C{a}@\rho_0$. Based on this
type template of $\C{data}$ field, we compute the class's template as
following:
\begin{center}
$\C{ListNode}\inang{\C{a}\extends\ObjZ}
\inang{\rhoalloc_0,\rho_0\,|\,\varphi_0} \extends \ObjZ \{...\}$
\end{center}
Next, we trivially templatize the type of $\C{next}$ field in the body
of the class as $\C{ListNode}\inang{\C{a}}\inang{\rhoalloc_0,\rho_0}$.
The resulting class represents a linked list with spine in the region
$\rhoalloc_0$ and data objects in the region $\rho_0$.

% Observe that our strategy of trivially templatizing the recursive
% occurances of a class type in its definition ensures that no new
% region variables are introduced.
The templatization technique we described above for recursive class
definitions can be extended to mutually recursive definitions in a
straightforward manner, by simultaneously templatizing them. Using the
techniques outlined above, we compute region type templates for all
classes bound in the class table of the $\absof{\FB}$ program before
we proceed to elaborate expressions.

@@

\subsection{Expression Elaboration}

Elaborating $\absof{\FB}$ expressions to $\FB$ expressions involves
(a). replacing core types in variable declarations and \C{new}
expressions with fresh region type templates, and (b). explicitly
instantiating region parameters of methods with fresh region variables
in method calls and function applications. This elaboration is
performed with respect to the polymorphic type templates of classes
and methods computed as per \S\ref{sec:fb-templatization}.

Function $\elabExpr$, shown in Fig.~\ref{fig:fb-elabexpr}, performs
this elaboration for a subset of expressions in $\absof{\FB}$, whose
corresponding $\FB$ expressions have been ascribed static semantics in
Fig.~\ref{fig:fb-staticsem}. $\elabExpr$ is defined under the same
context as the expression typing judgment in
Fig.~\ref{fig:fb-staticsem} with symbols $\A$,$\ralloc$, and $\env$
retaining their meaning. The function traverses expressions in a
syntax-directed manner of a type checker, introducing fresh region
type templates for unknown region types, while generating constraints
over region and predicate variables. The precise nature of these
constraints will be clarified shortly, but their essential nature is
that they capture the relationship between the type templates of
various subexpressions and their well-formedness. Note that
$\elabExpr$ returns the region type template of the subexpression,
which is used to generate constraints for the expression. Functions
$\typeOk$ and $\subtypeOk$ (definitions not shown) used by $\elabExpr$
implement type well-formedness and subtype judgments from
Fig.~\ref{fig:fb-staticsem}, respectively.

\subsection{Method and Class Elaboration}

Functions $\elabMeth$ and $\elabClass$ shown in
Fig.~\ref{fig:fb-elabmeth} lift expression elaboration to method and
class definitions, respectively. Both functions first build a context
($\A$) containing a set ($ \rhoenv$) of region variables denoting
regions that are currently live, a map ($\aenv$) mapping type
variables to their bounds, and a constraint formula ($\phicx$)
capturing constraints over live region variables. We use predicate
variables ($\varphi$ and $\varphi_m$) to capture constraints over
variables in $\rhoenv$ denoting the fact that such constraints are yet
to be inferred.

Function $\elabMeth$ elaborates a method definition of class $B$. It
calls $\elabExpr$ with the context $\A$, its allocation context
parameter ($\rhoallocm$), and a type environment ($\env$) that
contains region type bindings for all the arguments of the method,
including the implicit $\C{this}$ argument. The region type template
returned by $\elabExpr$ for the method body is checked against its
expected type (derived from the type template of the method)
generating more constraints. The function then returns the elaborated
method definition and the set of constraints.

$\elabClass$ elaborates a class's constructor ($k$) and method
definitions ($\bar{d}$) using $\elabCons$\footnote{The definition of
$\elabCons$ is straightforward, hence not shown.} and $\elabMeth$
functions, respectively. To the set of constraints returned by these
functions, $\elabClass$ adds constraints generated by checking the
well-formedness of the type templates of its superclass and fields,
and also a new constraint capturing the safety condition that
allocation regions of objects referred by the instance variables
should outlive the allocation region of the instance itself.

Function $\elabClassTable$ (Fig.~\ref{fig:fb-elabmeth}) elaborates
every definition in the class table $CT$, while accumulating
constraints. The constraints are finally solved by $\solve$
(\S\ref{sec:fb-constraintsolving}), which returns substitution
functions $\substFn_\rho$ and $\substFn_\varphi$ for free region and
predicate variables, respectively, introduced during templatization
and elaboration stages. The substitutions are applied to the class
table (and to the artifacts that make up the class table, recursively)
to compute a class table that maps classes to their fully
region-annotated definitions in $\FB$.

Note that if the original program in $\absof{\FB}$ contains unsafe
references, for example, a reference from a transferable region to a
stack regions, then the constraints generated during the elaboration
are not satisfiable. In such case, $\solve$ fails to solve
constraints, causing the program to be rejected.

\subsection{Constraint Solving}
\label{sec:fb-constraintsolving}

### 2015-11-02 paper/type-system.tex
Edits.

Class definitions in \FB extend Java's with region parameters
($\rho$). Informally, region variables ($\rho$) are to region
identifiers ($\rgn$) as type variables ($a$) are to core types ($T$);
they can be instantiated with region identifiers, as needed. A class
in \FB is necessarily parameterized over the allocation region of its
objects. We use $\rho^a$ to denote this parameter, and often refer to
it as \emph{class}'s allocation region, although it is really the
allocation region of its objects. An Object in \FB can contain fields
referring to objects allocated in regions ($\rhobar$) other than its
own allocation region ($\rhoalloc$), provided that the former outlive
the later (i.e., $\rhobar \outlives \rho$). In such cases, the
definition of object's class needs to be parametric over allocation
regions of its fields (i.e., their classes). Furthermore, the
constraint that such regions must outlive the allocation region of the
class needs to be made explicit in the definition. For instance, a
\C{Pair} object allocated in a region ${\rhoalloc}$ can have its first
component coming from ${\rho_1}$ and second from $\rho_2$, where all
the $\rho$'s are distinct, and $\rho_1, \rho_2$ outlive
$\rhoalloc$. The corresponding definition of the \C{Pair} class is as
follows(The symbol $\extends$ should be read \emph{extends}):  
\begin{codejava}[mathescape=true]
class Pair<a $\extends$ Object, b $\extends$ Object>
          <$\rho^a, \rho_1, \rho_2 \,|\, \rho_1 \outlives \rho^a \conj \rho_2 \outlives \rho^a$>
           {
  a@$\rho_1$ fst; 
  b@$\rho_2$ snd;
  Pair(a@$\rho_1$ fst, b@$\rho_2$ snd) {
    super(); 
    this.fst = fst; 
    this.snd = snd;
  }
  a@$\rho_1$ getFst() {
    return this.fst;
  }
}
\end{codejava}
To construct objects of the \C{Pair} class, its type and region
parameters need to be instantiated with core types ($T$) and concrete
region identifiers ($\rgn$), respectively. For example:
\begin{codejava}
letregion $\rgn_0$ {
  Object<$\rgn_0$> snd = new Object<$\rgn_0$>();
  letregion $\rgn_1$ {
    Object<$\rgn_1$> fst = new Object<$\rgn_1$>();
    let p = new Pair<Object,Object><$\rgn_1$,$\rgn_0$,$\rgn_1$>
                  (fst,snd);
  }
}
\end{codejava}
In the above code, the instantiation of $\rhoalloc$ and $\rho_1$ with
$\rgn_0$, and $\rho_2$ with $\rgn_1$ is allowed because (a) $\rgn_0$
and $\rgn_1$ are live during the instantation, and (b). $\rgn_0
\outlives \rgn_1$ and $\rgn_1 \outlives \rgn_1$ (since outlives is
reflexive). Observe that the region type of \C{p} conveys
the fact that (a). it is allocated in region $\rgn_1$, and (b). it
holds references to objects allocated in region $\rgn_0$ and $\rgn_1$.
In contrast, if we choose to allocate the \C{snd} object also in
$\rgn_1$, the region type of \C{p} would be
\C{Pair<\ObjZ,\ObjZ><$\rgn_1$,$\rgn_1$,$\rgn_1$>}, which we abbreviate
as \C{Pair<\ObjZ,\ObjZ>@$\rgn_1$}\footnote{In general, we treat
$B\inang{\tbar}@\rgn$ as an abbreviation of
$B\inang{\tbar}\inang{\rbar}$}. The $\ObjZ$ class in \FB has no
fields, hence its region type is of form $\ObjZ@\rgn$.
% In general, the \C{@} notation in a
% region type of an object \C{x} highlights that
% \C{x}, and all the objects reachable from \C{x} via references are
% allocated in a single region. We say that \C{x} is \emph{contained} in
% the region. The $\ObjZ$ class in \FB contains no references to other
% objects, hence its objects are always contained in their allocation
% region. Their region type is $\ObjZ\inang{\rgn}$ (or equivalently,
% $\ObjZ@\rgn$), for some region $\rgn$.

@@

\noindent The well-formedness judgment on region types
($\tywf{\A}{\tau}$) makes use of the well-formedness and subtyping
judgments on core types \footnote{We use a double-piped turnstile
($\Vdash$) for judgments in FGJ~\cite{fgj}, and a simple turnstile
($\vdash$) for those in \FB.}. All judgments are implicitly
parameterized on a class table ($CT \in cn \mapsto C$) that maps class
names to their definitions in \FB. For judgments in FGJ, we use a
class table that maps class names to their region-erased definitions.
We denote the region-erased class table as ($\absof{CT}$). The precise
semantics of region erasure ($\absof{\cdot}$) are defined
in~\cite{techrep}, but it can be understood as an operation that
erases region annotations on types. The class table is assumed to bind
the name $\RgnZ$ to the stub shown in Fig.~\ref{fig:region-stub}.
This mostly eliminates the need to treat $\RgnZ$ objects, which double
up as transferable region handlers, as a special case in semantics.

>>>>

Class definitions in \FB extend FGJ's with region parameters
($\rho$). Informally, region variables ($\rho$) are to region
identifiers ($\rgn$) as type variables ($a$) are to core types ($T$);
they can be instantiated with region identifiers, as needed. A class
in \FB is necessarily parameterized over the allocation region of its
objects. We use $\rho^a$ to denote this parameter, and often refer to
it as \emph{class}'s allocation region, although it is really the
allocation region of its objects. An Object in \FB can contain fields
referring to objects allocated in regions ($\rhobar$) other than its
own allocation region ($\rhoalloc$), provided that the former outlive
the later (i.e., $\rhobar \outlives \rho$). In such cases, the
definition of object's class needs to be parametric over allocation
regions of its fields (i.e., their classes). Furthermore, the
constraint that such regions must outlive the allocation region of the
class needs to be made explicit in the definition. For instance, a
\C{Pair} object allocated in a region ${\rhoalloc}$ can have its first
component coming from ${\rho_1}$ and second from $\rho_2$, where all
the $\rho$'s are distinct, and $\rho_1, \rho_2$ outlive
$\rhoalloc$. The corresponding definition of the \C{Pair} class is as
follows(The symbol $\extends$ should be read \emph{extends}):  
\begin{codejava}[mathescape=true]
class Pair<a $\extends$ Object, b $\extends$ Object>
          <$\rho^a, \rho_1, \rho_2 \,|\, \rho_1 \outlives \rho^a \conj \rho_2 \outlives \rho^a$>
           {
  a@$\rho_1$ fst; 
  b@$\rho_2$ snd;
  Pair(a@$\rho_1$ fst, b@$\rho_2$ snd) {
    super(); 
    this.fst = fst; 
    this.snd = snd;
  }
  a@$\rho_1$ getFst() {
    return this.fst;
  }
}
\end{codejava}
To construct objects of the \C{Pair} class, its type and region
parameters need to be instantiated with core types ($T$) and concrete
region identifiers ($\rgn$), respectively. For example:
\begin{codejava}
letregion $\rgn_0$ {
  Object<$\rgn_0$> snd = new Object<$\rgn_0$>();
  letregion $\rgn_1$ {
    Object<$\rgn_1$> fst = new Object<$\rgn_1$>();
    let p = new Pair<Object,Object><$\rgn_1$,$\rgn_0$,$\rgn_1$>
                  (fst,snd);
  }
}
\end{codejava}
In the above code, the instantiation of $\rhoalloc$ and $\rho_1$ with
$\rgn_0$, and $\rho_2$ with $\rgn_1$ is allowed because (a) $\rgn_0$
and $\rgn_1$ are live during the instantation, and (b). $\rgn_0
\outlives \rgn_1$ and $\rgn_1 \outlives \rgn_1$ (since outlives is
reflexive). Observe that the region type of \C{p} conveys
the fact that (a). it is allocated in region $\rgn_1$, and (b). it
holds references to objects allocated in region $\rgn_0$ and $\rgn_1$.
In contrast, if we choose to allocate the \C{snd} object also in
$\rgn_1$, the region type of \C{p} would be
\C{Pair<\ObjZ,\ObjZ><$\rgn_1$,$\rgn_1$,$\rgn_1$>}, which we abbreviate
as \C{Pair<\ObjZ,\ObjZ>@$\rgn_1$}\footnote{In general, we treat
$B\inang{\tbar}@\rgn$ as an abbreviation of
$B\inang{\tbar}\inang{\rbar}$}. The $\ObjZ$ class in \FB has no
fields, hence its region type is of form $\ObjZ@\rgn$.
% In general, the \C{@} notation in a
% region type of an object \C{x} highlights that
% \C{x}, and all the objects reachable from \C{x} via references are
% allocated in a single region. We say that \C{x} is \emph{contained} in
% the region. The $\ObjZ$ class in \FB contains no references to other
% objects, hence its objects are always contained in their allocation
% region. Their region type is $\ObjZ\inang{\rgn}$ (or equivalently,
% $\ObjZ@\rgn$), for some region $\rgn$.

@@

\noindent The well-formedness judgment on region types
($\tywf{\A}{\tau}$) makes use of the well-formedness and subtyping
judgments on core types \footnote{We use a double-piped turnstile
($\Vdash$) for judgments in FGJ~\cite{fgj}, and a simple turnstile
($\vdash$) for those in \FB.}. All judgments are implicitly
parameterized on a class table ($CT \in cn \mapsto D$) that maps class
names to their definitions in \FB. For judgments in FGJ, we use a
class table that maps class names to their region-erased definitions.
We denote the region-erased class table as ($\absof{CT}$). The precise
semantics of region erasure ($\absof{\cdot}$) are defined
in~\cite{techrep}, but it can be understood as an operation that
erases region annotations on types. The class table is assumed to bind
the name $\RgnZ$ to the stub shown in Fig.~\ref{fig:region-stub}.
This mostly eliminates the need to treat $\RgnZ$ objects, which double
up as transferable region handlers, as a special case in semantics.
